\chapter{Tensor Product Space Transformations}
Let's consider our good ol' vector space $V$ with basis $\{\vec{e}_\mu\}$ and its dual space $V^*$ with basis $\{\vec{e}^\nu\}$.
We now know that $\{\vec{e}_\mu\}$ will transform covariantly, while $\{\vec{e}^\nu\}$ will transform contravariantly.
We always choose $\{e^\nu\}$ such that $\langle \vec{e}^\nu, \vec{e}_\mu \rangle = \delta^\nu_\mu$.
Now that we've learned about change of basis transformations, let's construct alternate bases for $V$ and $V^*$, called $\{\unit{e}_\mu\}$ and $\{\unit{e}^\nu\}$, again with the property that $\langle \unit{e}_\nu, \unit{e}_\mu \rangle = \delta_\mu^\nu$.
Let's also say for the sake of argument that $\{\unit{e}_\mu\}$ transforms into $\{\vec{e}_\mu\}$ using a transformation $\Lambda$, so we then have that
\[ \unit{e}_\nu = \tensor{\Lambda}{_\nu^\mu} \vec{e}_\mu. \]
Because the bases of the dual space are contravariant, we know that the equivalent transformation for $\{\unit{e}^\nu\}$ and $\{\vec{e}^\nu\}$ would be
\[ \unit{e}^\nu = \tensor{\qty(\Lambda^{-1})}{_\mu^\nu} \vec{e}^\mu. \]

To understand how transformations affect tensor product spaces, let's construct a rank $(1,2)$ tensor,
\[ \tensor{T}{^\lambda_{\mu\nu}} \vec{e}_\lambda \otimes \vec{e}^\mu \otimes \vec{e}^\nu, \]
and transform it from the $\vec{e}$ basis into the $\unit{e}$ basis.
This goes in the direction $\vec{e} \to \unit{e}$, which is the opposite way we've gone previously, but the technique remains the exact same.
We start by defining the transformation $\Lambda$ such that
\[ \vec{e}_\mu = \tensor{\Lambda}{_\mu^\nu} \unit{e}_\nu. \]
We then rewrite our original tensor definition, replace the original bases with their hatted variants and adding the appropriate variant of $\Lambda$, so that
\[ \tensor{T}{^\lambda_{\mu\nu}}\vec{e}_\lambda \otimes \vec{e}^\mu \otimes \vec{e}^\nu = \tensor{T}{^\lambda_{\mu\nu}} \Big[\tensor{\Lambda}{_\lambda^\alpha}\unit{e}_\alpha\Big] \otimes \qty[\tensor{\qty(\Lambda^{-1})}{_\beta^\mu}\unit{e}^\beta] \otimes \qty[\tensor{\qty(\Lambda^{-1})}{_\gamma^\nu}\unit{e}^\gamma]. \]
Because tensors are multilinear products, we can factor the $\Lambda$ and $\Lambda^{-1}$ out in front of our new bases to get
\[ \tensor{T}{^\lambda_{\mu\nu}}\Lambda\Lambda^{-1}\Lambda^{-1} \unit{e}_\lambda \otimes \unit{e}^\mu \otimes \unit{e}^\nu. \]
This means that our component $\tensor{\widehat{T}}{^\lambda_{\mu\nu}} = \tensor{T}{^\lambda_{\mu\nu}}\Lambda\Lambda^{-1}\Lambda^{-1}$.
This represents the transformation of the basis of the tensor; if we want to transform the component $\tensor{T}{^\lambda_{\mu\nu}}$, we need to use the opposite transformations, since each index of the component is opposite the index on the corresponding basis vector.
This means that 
\[ \tensor{T}{^\lambda_{\mu\nu}} = \tensor{\widehat{T}}{^\lambda_{\mu\nu}}\Lambda\Lambda\Lambda^{-1}. \]
Note that not only are the transformations inverted, but their order is also reversed; to understand why this is, let's examine the full transformation of our tensor from $\vec{e}$ to $\unit{e}$.
By replacing each piece of our original tensor with its transformed variant, we see that
\begin{align*}
    \tensor{T}{^\lambda_{\mu\nu}}\vec{e}_\lambda \otimes \vec{e}^\mu \otimes \vec{e}^\nu &= 
    \tensor{\widehat{T}}{^\lambda_{\mu\nu}}\Lambda\Lambda\Lambda^{-1}\Lambda\Lambda^{-1}\Lambda^{-1} \unit{e}_\lambda \otimes \unit{e}^\mu \otimes \unit{e}^\nu \\
    &= \tensor{\widehat{T}}{^\lambda_{\mu\nu}}\unit{e}_\lambda \otimes \unit{e}^\mu \otimes \unit{e}^\nu.
\end{align*}
Just like how using both $\Lambda$ and $\Lambda^{-1}$ to transform an arbitrary vector causes the transformations to cancel, reversing the order of $\Lambda$ when transforming the basis and the components does the same.

\subsection{Covariant and Contravariant Tensors}
The previous example discussed the transformation of a mixed, rank $(1,2)$ tensor.
Because the basis vectors of the tensor product space included both vectors and covectors, some parts of the tensor transformed covariantly and some transformed contravariantly.
This needn't always be the case though; consider a rank $(0,3)$ tensor, whose basis would be of the form $\vec{e}^\lambda \otimes \vec{e}^\mu \otimes \vec{e}^\nu$, and whose component would be $T_{\lambda\mu\nu}$.
This tensor has a basis comprised solely of covectors.
It's basis is then completely contravariant, and its component is entirely covariant.
This is a \emph{covariant tensor}.
Note that we always identify a tensors variance by the variance of its component, not its basis.
A contravariant tensor might be a rank $(3,0)$ tensor $T^{\lambda\mu\nu}$.
\begin{itemize}
    \item Covariant tensors are identified with all lower indices, like $T_{\lambda\mu\nu}$, and are of rank $(0,n)$. Their basis is comprised entirely of covectors, and they map $n$-tuples of vectors to a scalar field.
    \item Contravariant tensors have all upper indices, like $T^{\lambda\mu\nu}$, and are of rank $(n,0)$. Their basis is entirely products of vectors, and they map $n$-tuples of covectors to a scalar field.
\end{itemize}
